---
title: "SFER & MIR Habitat Context Maps"
author: "Heidi K. Hirsh"
date: "`r Sys.Date()`"

output:
  html_document:
    toc: true
    toc_float: true
    number_sections: false
    theme: flatly
    self_contained: false
output_dir: "docs"
output_file: "index.html"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,       # HIDE CODE
  message = FALSE,
  warning = FALSE,
  fig.width = 8.4,
  fig.height = 7.4,
  dpi = 300,
  comment = NA 
)

sf::sf_use_s2(FALSE)
```

<link rel="stylesheet" href="https://cdn.jsdelivr.net/npm/glightbox/dist/css/glightbox.min.css">
<script src="https://cdn.jsdelivr.net/npm/glightbox/dist/js/glightbox.min.js"></script>
<script>
document.addEventListener("DOMContentLoaded", function() {
  GLightbox({ selector: ".glightbox" });
});
</script>




<style>
/* --- Make tabsets vertical (left column) --- */
.tabset .nav-tabs,
.tabset .nav-pills {
  float: left;
  width: 220px;
  margin-right: 20px;
  border-bottom: none;
}

.tabset .nav-tabs > li,
.tabset .nav-pills > li {
  float: none;
  width: 100%;
}

.tabset .tab-content {
  overflow: auto; /* lets content sit to the right of the vertical tabs */
}

/* Keep the tab content from jumping weirdly */
.tabset .tab-content > .tab-pane {
  padding-top: 10px;
}

/* Optional: slightly smaller tab font */
.tabset .nav > li > a {
  padding: 8px 10px;
}
</style>


## Overview

This report generates **site-specific habitat context maps** for SFER monitoring stations and Mission: Iconic Reefs (MIR) sites in the Florida Keys. For each focal location, benthic habitat polygons, nearby SFER stations, and MIR site boundaries are visualized within a defined spatial context, using consistent symbology and labeling across all map products.

Maps are produced in four batches:

- **Batch A — SFER inshore (regional context)**  
  Habitat within **50 km** of each inshore SFER station, shown within a **30 km** map context.

- **Batch B — SFER inshore (zoomed)**  
  Habitat within **25 km** of each inshore SFER station, shown within a **10 km** map context.

- **Batch C1 — MIR sites (local context)**  
  Habitat within **10 km** of each MIR site centroid, shown within a **5 km** map context.  
  Focal MIR points are not drawn; site polygons and neighboring features provide spatial context.

- **Batch C2 — MIR sites (expanded context)**  
  Habitat within **10 km** of each MIR site centroid, shown within a **10 km** map context.

All maps include:
- Unified Florida Reef Map habitat polygons clipped to the habitat radius
- Florida Keys shoreline
- Contextual SFER stations and MIR site polygons
- A Keys-wide inset map indicating the focal map extent

Outputs are saved to a timestamped run directory under  
`Outputs_sfer_habitat_maps/`, with figures organized by batch and a combined CSV log summarizing map outputs and feature counts.


```{r libraries, include=FALSE}
suppressPackageStartupMessages({
  library(sf)
  library(dplyr)
  library(DT)
  library(ggplot2)
  library(leaflet)
  library(htmltools)
  library(readr)
  library(stringr)
  library(cowplot)
  library(ggrepel)
  library(grid)     # for unit()
  library(scales)   # for alpha() 
})
```

```{r settings, include=FALSE}
crs_metric <- 32617
inset_pad_frac <- 0.20

regen_maps <- FALSE   # <- set TRUE only when you want to re-generate PNGs

# Paths
shore_path <- "data/Florida_Shoreline_(1_to_12%2C000_Scale)/Florida_Shoreline_(1_to_12%2C000_Scale).shp"
hab_gdb     <- "data/FWC_UnifiedFloridaReefMap_v2.0.gdb"
hab_layer   <- "UnifiedReefMap"
sfer_path   <- "data/SFER_station_coordinates.csv"
mir_path    <- "data/MIR_Shapefiles/Overall Site/MIR Site Polygons.shp"

# SFER columns (match your file)
sfer_lon_col  <- "lon"
sfer_lat_col  <- "lat"
sfer_id_col   <- "SiteID"
sfer_zone_col <- "Zone"
inshore_value <- "Inshore"

# --------------------------------------------------
# Run folder control
# --------------------------------------------------
regen_maps <- FALSE   # TRUE = regenerate maps; FALSE = reuse existing PNGs

if (regen_maps) {
  ts <- format(Sys.time(), "%Y%m%d_%H%M")
  run_dir <- file.path("Outputs_sfer_habitat_maps", paste0("run_", ts))
  dir.create(file.path(run_dir, "Figures"), recursive = TRUE, showWarnings = FALSE)
  dir.create(file.path(run_dir, "Tables"),  recursive = TRUE, showWarnings = FALSE)
} else {
  # 👇 CHANGE THIS to the run folder that already exists on your machine
  # run_dir <- "Outputs_sfer_habitat_maps/run_20260205_1356"
    run_dir <- "docs/Outputs_sfer_habitat_maps/run_SHARE"


}

message("Using run directory: ", normalizePath(run_dir, winslash = "/", mustWork = FALSE))
```




```{r github-pages-paths, include=FALSE}
# Convert filesystem paths under docs/ to web-relative paths for GitHub Pages
to_web <- function(x) sub("^docs/", "", x)
```

```{r styling, include=FALSE}
# -----------------------------
# Styling (points, labels, inset)
# -----------------------------

# Point styling
pt_nonfocal_size   <- 2.6
pt_focal_size      <- 4.2
pt_stroke          <- 0.9
pt_nonfocal_fill   <- "black"
pt_nonfocal_border <- "white"
pt_focal_fill      <- "blue3"
pt_focal_border    <- "white"

# Label styling
label_size_main    <- 6.0   # bigger than before
label_size_small   <- 5.2
label_box_alpha    <- 0.55  # semi-transparent label background
label_pad          <- 0.20

# Inset placement
inset_x <- 0.70
inset_y <- 0.06
inset_w <- 0.24
inset_h <- 0.24

# Scale bar overlay placement (above inset)
# (kept, but no longer used)
scalebar_x <- inset_x
scalebar_y <- inset_y + inset_h + 0.01
scalebar_w <- inset_w
scalebar_h <- 0.06
```

```{r helper-expand-bbox, include=FALSE}
expand_bbox <- function(bb, expand_frac = 0.15) {
  xr <- as.numeric(bb["xmax"] - bb["xmin"])
  yr <- as.numeric(bb["ymax"] - bb["ymin"])
  bb["xmin"] <- bb["xmin"] - xr * expand_frac
  bb["xmax"] <- bb["xmax"] + xr * expand_frac
  bb["ymin"] <- bb["ymin"] - yr * expand_frac
  bb["ymax"] <- bb["ymax"] + yr * expand_frac
  bb
}
```

```{r helper-label-df, include=FALSE}
# -----------------------------
# Helper: make label df from sf points
# -----------------------------
sf_points_to_label_df <- function(x_sf, label_col) {
  if (is.null(x_sf) || nrow(x_sf) == 0) return(NULL)
  
  xy <- sf::st_coordinates(x_sf)
  
  out <- x_sf %>%
    sf::st_drop_geometry() %>%
    dplyr::mutate(
      .x   = xy[, 1],
      .y   = xy[, 2],
      .lbl = as.character(.data[[label_col]])
    )
  
  out
}

zone_cols <- c(
  "Inshore" = "#1f78b4",
  "Mid"     = "#33a02c",
  "Offshore"= "#e31a1c"
)
```

```{r helper-order, include=FALSE}
sfer_order <- c("1","EK_IN","5","4","UK_IN","7","10","13","16","19","24")  # <-- your SiteIDs in the order you want
mir_order  <- c("Carysfort_North","Carysfort_South","Horseshoe","Cheeca","Sombrero","NFH","Looe","EDR")

mir_name_lookup <- c(
  "EDR" = "Eastern Dry Rocks",
  "NFH" = "Newfound Harbor"
)

```

```{r load-data, include=FALSE}
FLKs1 <- st_read(shore_path, quiet = TRUE)
coral_sf <- st_read(dsn = hab_gdb, layer = hab_layer, quiet = TRUE)

sfer_df <- read_csv(sfer_path, show_col_types = FALSE)
stopifnot(all(c(sfer_lon_col, sfer_lat_col, sfer_id_col, sfer_zone_col) %in% names(sfer_df)))

sfer_df[[sfer_lon_col]] <- as.numeric(sfer_df[[sfer_lon_col]])
sfer_df[[sfer_lat_col]] <- as.numeric(sfer_df[[sfer_lat_col]])

sfer_sf <- st_as_sf(
  sfer_df,
  coords = c(sfer_lon_col, sfer_lat_col),
  crs = 4326,
  remove = FALSE
)

mir_sf <- NULL
if (file.exists(mir_path)) {
  mir_sf <- st_read(mir_path, quiet = TRUE) %>% st_make_valid()
} else {
  message("MIR shapefile not found; MIR loop will be skipped: ", mir_path)
}
```

```{r habitat-palette, include=FALSE}
habs1 <- c(
  "Aggregate Reef","Individual or Aggregated Patch Reef",
  "Reef Rubble","Ridge",
  "Pavement",
  "Scattered Coral/Rock in Unconsolidated Sediment","Unconsolidated Sediment",
  "Seagrass (Continuous)","Seagrass (Discontinuous)",
  "Mangrove",
  "Land","Artificial","Dredged/Excavated",
  "Not Classified"
)

class1_colors <- c(
  "maroon1","red",
  "blueviolet","mediumturquoise",
  "steelblue1",
  "coral","khaki",
  "darkgreen","forestgreen",
  "saddlebrown",
  "gray50","gray50","navyblue",
  "white"
)

hab_cols <- setNames(class1_colors, habs1)

class1_FILL_scale <- scale_fill_manual(
  name   = "Class 1 Habitat",
  values = hab_cols,
  breaks = habs1,
  labels = habs1,
  drop   = FALSE
)
```

```{r transforms-and-inset, include=FALSE}
FLKs1_ll <- st_transform(FLKs1, 4326)
coral_ll <- st_transform(st_make_valid(coral_sf), 4326)

sfer_all_ll <- st_transform(sfer_sf, 4326)
sfer_all_m  <- st_transform(sfer_sf, crs_metric)

sfer_in <- sfer_sf %>%
  filter(str_trim(str_to_lower(.data[[sfer_zone_col]])) == str_to_lower(inshore_value))
if (nrow(sfer_in) == 0) stop("No inshore SFER stations found.")

sfer_in_m <- st_transform(sfer_in, crs_metric)

mir_all_ll <- if (!is.null(mir_sf)) st_transform(mir_sf, 4326) else NULL
mir_all_m  <- if (!is.null(mir_sf)) st_transform(mir_sf, crs_metric) else NULL

bb_inset <- st_bbox(sfer_all_ll)
if (!is.null(mir_all_ll)) {
  bb_mir <- st_bbox(mir_all_ll)
  bb_inset["xmin"] <- min(bb_inset["xmin"], bb_mir["xmin"])
  bb_inset["ymin"] <- min(bb_inset["ymin"], bb_mir["ymin"])
  bb_inset["xmax"] <- max(bb_inset["xmax"], bb_mir["xmax"])
  bb_inset["ymax"] <- max(bb_inset["ymax"], bb_mir["ymax"])
}
bb_inset <- expand_bbox(bb_inset, inset_pad_frac)

FLKs1_inset <- st_crop(FLKs1_ll, bb_inset)
mir_inset   <- if (!is.null(mir_all_ll)) st_crop(mir_all_ll, bb_inset) else NULL

p_inset_base <- ggplot() +
  geom_sf(data = FLKs1_inset, fill = "gray50", color = NA) +
  { if (!is.null(mir_inset)) geom_sf(data = mir_inset, fill = NA, color = "orange", linewidth = 0.25, alpha = 0.8) else NULL } +
  geom_sf(data = sfer_all_ll, color = "black", size = 0.55, alpha = 0.6) +
  coord_sf(
    xlim = c(bb_inset["xmin"], bb_inset["xmax"]),
    ylim = c(bb_inset["ymin"], bb_inset["ymax"]),
    expand = FALSE
  ) +
  theme_void() +
  theme(plot.background = element_rect(fill = "white", color = "gray40", linewidth = 0.2))
```

## Reference Map
```{r reference-map-static, echo=FALSE, results='asis'}
# ============================
# Reference Map (STATIC, Keys-only)
# - ALWAYS regenerates (no caching)
# - Default ggplot colors by Zone column
# - SFER labels small + black inside points (no leader lines)
# - MIR polygons thin black outline
# - MIR labels allowed to have leader lines + pushed farther away
# ============================

# Output image path (overwrites every knit)
ref_png <- file.path(run_dir, "Figures", "reference_map_static_keys.png")

# ----------------------------
# 1) Build Keys-only extent from data (SFER + MIR)
# ----------------------------
bb_keys <- sf::st_bbox(sfer_all_ll)

if (!is.null(mir_all_ll) && nrow(mir_all_ll) > 0) {
  bb_m <- sf::st_bbox(mir_all_ll)
  bb_keys["xmin"] <- min(bb_keys["xmin"], bb_m["xmin"])
  bb_keys["ymin"] <- min(bb_keys["ymin"], bb_m["ymin"])
  bb_keys["xmax"] <- max(bb_keys["xmax"], bb_m["xmax"])
  bb_keys["ymax"] <- max(bb_keys["ymax"], bb_m["ymax"])
}

# Pad the bbox a bit so labels have room (simple + stable; no NA bbox math)
pad_x <- as.numeric(bb_keys["xmax"] - bb_keys["xmin"]) * 0.10
pad_y <- as.numeric(bb_keys["ymax"] - bb_keys["ymin"]) * 0.15
bb_keys["xmin"] <- bb_keys["xmin"] - pad_x
bb_keys["xmax"] <- bb_keys["xmax"] + pad_x
bb_keys["ymin"] <- bb_keys["ymin"] - pad_y
bb_keys["ymax"] <- bb_keys["ymax"] + pad_y

# Crop geometry using bbox (robust way)
bb_sfc <- sf::st_as_sfc(bb_keys)

FLKs_keys <- sf::st_crop(FLKs1_ll, bb_sfc)
sfer_keys <- sf::st_crop(sfer_all_ll, bb_sfc)
mir_keys  <- if (!is.null(mir_all_ll) && nrow(mir_all_ll) > 0) sf::st_crop(mir_all_ll, bb_sfc) else NULL

# ----------------------------
# 2) MIR label positions (for polygons)
# ----------------------------
mir_lab_df <- NULL
if (!is.null(mir_keys) && nrow(mir_keys) > 0 && "MIRSite" %in% names(mir_keys)) {
  mir_pts <- sf::st_point_on_surface(mir_keys)
  mir_lab_df <- sf_points_to_label_df(mir_pts, "MIRSite")
}

# ----------------------------
# 3) Plot
# ----------------------------
p_ref <- ggplot() +

  # Shoreline
  geom_sf(
    data = FLKs_keys,
    fill = "gray85",
    color = NA
  ) +

  # MIR polygons — thin outline
  { if (!is.null(mir_keys) && nrow(mir_keys) > 0)
    geom_sf(
      data = mir_keys,
      fill = NA,
      color = "black",
      linewidth = 0.25,
      alpha = 0.9
    )
    else NULL } +

  # MIR labels — leader lines OK, pushed farther away
  { if (!is.null(mir_lab_df) && nrow(mir_lab_df) > 0)
    ggrepel::geom_label_repel(
      data = mir_lab_df,
      aes(x = .x, y = .y, label = .lbl),
      size = 3.2 / ggplot2::.pt,
      fontface = "bold",
      label.size = 0,
      fill = scales::alpha("white", 0.70),
      box.padding = 0.7,
      point.padding = 0.6,
      force = 2.5,
      min.segment.length = 0,
      segment.size = 0.25,
      max.overlaps = Inf
    )
    else NULL } +

  # SFER points — default ggplot colors by Zone (no manual palette)
  geom_sf(
    data = sfer_keys,
    aes(fill = .data[[sfer_zone_col]]),
    shape = 21,
    color = "white",
    stroke = 0.6,
    size = 4.0,
    alpha = 0.95
  ) +

  # SFER labels — small, black, inside points (no leader lines)
  geom_sf_text(
    data = sfer_keys,
    aes(label = .data[[sfer_id_col]]),
    color = "black",
    size = 1.8,
    fontface = "plain"
  ) +

  coord_sf(
    xlim = c(bb_keys["xmin"], bb_keys["xmax"]),
    ylim = c(bb_keys["ymin"], bb_keys["ymax"]),
    expand = FALSE
  ) +

  theme_bw() +
  theme(
    panel.grid = element_blank(),
    legend.title = element_text(size = 10),
    legend.text  = element_text(size = 9)
  ) +
  labs(
    fill  = "SFER Zone",
    title = "SFER stations and MIR sites — Florida Keys"
  )

# ALWAYS save (overwrite)
ggsave(ref_png, p_ref, width = 9.5, height = 6.5, dpi = 300)

# Display with click-to-zoom
cat(sprintf(
  "<a href='%s' class='glightbox' data-gallery='reference'>
     <img src='%s' style='width:100%%; max-width:1100px; height:auto;
          border:1px solid #ddd; border-radius:6px;'/>
   </a>",
  ref_png, ref_png
))


```

```{r make-maps-function, include=FALSE}
make_maps <- function(
    focal_ll, focal_m, focal_id_vec,
    id_name = "ID",
    out_subdir = "Batch",
    hab_km = 40, context_km = 30,
    include_inset = TRUE,
    force_single_point_per_id = TRUE,
    label_sfer_points = TRUE,
    label_mir_sites   = TRUE,
    mir_name_field    = "MIRSite",
    draw_focal_point  = TRUE 
) {
  
  fig_dir <- file.path(run_dir, "Figures", out_subdir)
  dir.create(fig_dir, recursive = TRUE, showWarnings = FALSE)
  
  hab_m <- hab_km * 1000
  ctx_m <- context_km * 1000
  
  # Preclip habitat to union of buffers sized to max(hab, ctx)
  preclip_m <- max(hab_m, ctx_m)
  buf_union_pre <- st_union(st_buffer(focal_m, dist = preclip_m))
  coral_pre_m <- st_transform(coral_ll, crs_metric) %>% st_intersection(buf_union_pre)
  
  out_log <- list()
  
  for (sid in focal_id_vec) {
    
    message("[", out_subdir, "] focal ", id_name, ": ", sid)
    
    pt_ll <- focal_ll %>% filter(.data[[id_name]] == sid)
    pt_m  <- focal_m  %>% filter(.data[[id_name]] == sid)
    
    if (force_single_point_per_id && nrow(pt_m) > 1) {
      pt_m  <- pt_m  %>% slice(1)
      pt_ll <- pt_ll %>% slice(1)
    }
    
    pt_u_m <- st_union(pt_m)
    
    buf_ctx_m <- st_buffer(pt_u_m, dist = ctx_m)
    buf_hab_m <- st_buffer(pt_u_m, dist = hab_m)
    
    buf_ctx_ll <- st_transform(buf_ctx_m, 4326)
    buf_hab_ll <- st_transform(buf_hab_m, 4326)
    
    coral_clip_ll <- st_intersection(st_transform(coral_pre_m, 4326), buf_hab_ll)
    
    # context SFER
    sfer_ctx_m  <- sfer_all_m[st_intersects(sfer_all_m, buf_ctx_m, sparse = FALSE), ]
    sfer_ctx_ll <- st_transform(sfer_ctx_m, 4326)
    
    # drop focal from "context" so it can't be drawn/labeled twice
    sfer_ctx_ll_nonfocal <- sfer_ctx_ll %>%
      filter(.data[[sfer_id_col]] != sid)
    
    # MIR context
    mir_ctx_ll <- NULL
    n_mir <- NA_integer_
    if (!is.null(mir_all_m)) {
      mir_ctx_m <- mir_all_m[st_intersects(mir_all_m, buf_ctx_m, sparse = FALSE), ]
      n_mir <- nrow(mir_ctx_m)
      if (n_mir > 0) mir_ctx_ll <- st_transform(mir_ctx_m, 4326)
    }
    
    # If focal IDs are MIRSite names, drop focal MIR polygon from ctx labels to prevent duplicates
    mir_ctx_ll_nonfocal <- mir_ctx_ll
    if (!is.null(mir_ctx_ll) && mir_name_field %in% names(mir_ctx_ll)) {
      mir_ctx_ll_nonfocal <- mir_ctx_ll %>% filter(.data[[mir_name_field]] != sid)
    }
    
    # bbox extent = context buffer
    bb_main <- st_bbox(buf_ctx_ll)
    bb_rect <- st_as_sfc(bb_main) %>% st_sf(geometry = ., crs = 4326)
    
    # Label dfs
    sfer_lab_df_nonfocal <- if (label_sfer_points) sf_points_to_label_df(sfer_ctx_ll_nonfocal, sfer_id_col) else NULL
    sfer_lab_df_focal    <- if (label_sfer_points) sf_points_to_label_df(pt_ll %>% mutate(tmpID = sid), "tmpID") else NULL
    
    mir_lab_df_nonfocal  <- NULL
    if (label_mir_sites && !is.null(mir_ctx_ll_nonfocal) && nrow(mir_ctx_ll_nonfocal) > 0 && mir_name_field %in% names(mir_ctx_ll_nonfocal)) {
      mir_pts <- st_point_on_surface(mir_ctx_ll_nonfocal)
      mir_lab_df_nonfocal <- sf_points_to_label_df(mir_pts, mir_name_field)
    }
    
    # ---- Title prefix (MUST be defined outside ggplot chain) ----
    title_prefix <- if (grepl("^[0-9]+$", as.character(sid))) "Site " else "MIR "
    
    # MAIN plot
    p_main <- ggplot(data = FLKs1_ll) +
      { if (nrow(coral_clip_ll) > 0)
        geom_sf(data = coral_clip_ll, aes(fill = ClassLv1), linewidth = 0, alpha = 0.7)
        else NULL } +
      geom_sf(fill = "gray50", linewidth = 0) +
      
      # MIR polygons (context)
      { if (!is.null(mir_ctx_ll) && nrow(mir_ctx_ll) > 0)
        geom_sf(data = mir_ctx_ll, fill = NA, color = "black", linewidth = 0.6, alpha = 0.9)
        else NULL } +
      
      # SFER points (context non-focal)
      geom_sf(
        data = sfer_ctx_ll_nonfocal,
        shape = 21, fill = pt_nonfocal_fill, color = pt_nonfocal_border,
        size = pt_nonfocal_size, stroke = pt_stroke, alpha = 0.95
      ) +

      
      { if (draw_focal_point)
        geom_sf(
          data = pt_ll,
          shape = 21, fill = pt_focal_fill, color = pt_focal_border,
          size = pt_focal_size, stroke = pt_stroke, alpha = 1
        )
        else NULL }+
      
      # Labels: SFER non-focal
      { if (!is.null(sfer_lab_df_nonfocal) && nrow(sfer_lab_df_nonfocal) > 0)
        geom_label_repel(
          data = sfer_lab_df_nonfocal,
          aes(x = .x, y = .y, label = .lbl),
          size = label_size_small / ggplot2::.pt,
          label.size = 0,
          fill = scales::alpha("white", label_box_alpha),
          label.padding = unit(label_pad, "lines"),
          min.segment.length = 0,
          segment.size = 0.25,
          max.overlaps = 50
        )
        else NULL } +
      
      # Label: focal (ONLY ONCE)
      { if (!is.null(sfer_lab_df_focal) && nrow(sfer_lab_df_focal) > 0)
        geom_label_repel(
          data = sfer_lab_df_focal,
          aes(x = .x, y = .y, label = .lbl),
          fontface = "bold",
          size = label_size_main / ggplot2::.pt,
          label.size = 0,
          fill = scales::alpha("white", label_box_alpha),
          label.padding = unit(label_pad, "lines"),
          min.segment.length = 0,
          segment.size = 0.35,
          max.overlaps = Inf
        )
        else NULL } +
      
      # Labels: MIR non-focal polygons
      { if (!is.null(mir_lab_df_nonfocal) && nrow(mir_lab_df_nonfocal) > 0)
        geom_label_repel(
          data = mir_lab_df_nonfocal,
          aes(x = .x, y = .y, label = .lbl),
          fontface = "bold",
          size = label_size_small / ggplot2::.pt,
          label.size = 0,
          fill = scales::alpha("white", label_box_alpha),
          label.padding = unit(label_pad, "lines"),
          min.segment.length = 0,
          segment.size = 0.25,
          max.overlaps = Inf
        )
        else NULL } +
      
      class1_FILL_scale +
      coord_sf(
        xlim = c(bb_main["xmin"], bb_main["xmax"]),
        ylim = c(bb_main["ymin"], bb_main["ymax"]),
        expand = FALSE
      ) +
      theme_bw() +
      theme(
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text  = element_text(size = 12),
        axis.title = element_text(size = 14),
        legend.position = "right",
        legend.title = element_text(size = 10),
        legend.text  = element_text(size = 9)
      ) +
      xlab("Longitude") + ylab("Latitude") +
      # ggtitle(paste0(as.character(sid), " (context ", context_km, " km)"))
      ggtitle(paste0(title_prefix, as.character(sid), " (scale ", context_km, " km)"))
    
    
    
    # Inset (scale bar overlay REMOVED)
    if (include_inset) {
      
      p_inset <- p_inset_base +
        geom_sf(data = bb_rect, fill = NA, color = "red", linewidth = 0.6) +
        { if (draw_focal_point)
          geom_sf(
            data = pt_ll,
            shape = 21, fill = pt_focal_fill, color = pt_focal_border,
            size = 2.2, stroke = 0.8
          )
          else NULL }
      
      p <- ggdraw() +
        draw_plot(p_main, 0, 0, 1, 1) +
        # draw_plot(p_scale, x = scalebar_x, y = scalebar_y, width = scalebar_w, height = scalebar_h) +  # REMOVED
        draw_plot(p_inset, x = inset_x, y = inset_y, width = inset_w, height = inset_h)
      
    } else {
      p <- p_main
    }
    
    safe_id <- str_replace_all(as.character(sid), "[^A-Za-z0-9]+", "_")
    out_png <- file.path(fig_dir, paste0(out_subdir, "_hab", hab_km, "km_ctx", context_km, "km_", safe_id, ".png"))
    
    ggsave(out_png,p,width = 8.4,height = 7.4,dpi = 300,create.dir = TRUE)
    
    out_log[[length(out_log) + 1]] <- data.frame(
      batch = out_subdir,
      focal_id = as.character(sid),
      hab_km = hab_km,
      context_km = context_km,
      n_hab_polys = nrow(coral_clip_ll),
      n_sfer_in_context = nrow(sfer_ctx_ll),
      n_mir_in_context = n_mir,
      file = out_png
    )
  }
  
  bind_rows(out_log)
}
```

```{r run-batches, include=FALSE, eval=regen_maps, message=TRUE}
# -----------------------------
# 6) Batch A: SFER inshore
# -----------------------------
sfer_focal_ll <- sfer_in %>% mutate(ID = .data[[sfer_id_col]])
sfer_focal_m  <- sfer_in_m %>% mutate(ID = .data[[sfer_id_col]])
sfer_ids <- unique(sfer_focal_ll$ID)

log_A <- make_maps(
  focal_ll = sfer_focal_ll, focal_m = sfer_focal_m, focal_id_vec = sfer_ids,
  id_name = "ID",
  out_subdir = "SFER_context30_hab50",
  hab_km = 50, context_km = 30,
  include_inset = TRUE,
  label_sfer_points = TRUE,
  label_mir_sites   = TRUE,
  mir_name_field    = "MIRSite",
  draw_focal_point  = TRUE
)

# -----------------------------
# 7) Batch B: SFER inshore ZOOM
# -----------------------------
log_B <- make_maps(
  focal_ll = sfer_focal_ll, focal_m = sfer_focal_m, focal_id_vec = sfer_ids,
  id_name = "ID",
  out_subdir = "SFER_context10_hab25_zoom",
  hab_km = 25, context_km = 10,
  include_inset = TRUE,
  label_sfer_points = TRUE,
  label_mir_sites   = TRUE,
  mir_name_field    = "MIRSite",
  draw_focal_point  = TRUE
)

# -----------------------------
# 8) Batch C: MIR sites (use mir_sf$MIRSite)
# -----------------------------
log_C  <- NULL
log_C2 <- NULL

if (!is.null(mir_sf)) {

  if (!("MIRSite" %in% names(mir_sf))) {
    stop("Expected mir_sf$MIRSite but did not find column 'MIRSite' in MIR shapefile.")
  }

  mir_m <- st_transform(mir_sf, crs_metric)
  mir_pts_m <- st_point_on_surface(mir_m) %>% st_sf() %>%
    mutate(ID = as.character(.data[["MIRSite"]]))

  mir_pts_ll <- st_transform(mir_pts_m, 4326)
  mir_ids <- unique(mir_pts_ll$ID)

  # C1
  log_C <- make_maps(
    focal_ll = mir_pts_ll, focal_m = mir_pts_m, focal_id_vec = mir_ids,
    id_name = "ID",
    out_subdir = "MIR_context5_hab10",
    hab_km = 10, context_km = 5,
    include_inset = TRUE,
    label_sfer_points = TRUE,
    label_mir_sites   = TRUE,
    mir_name_field    = "MIRSite",
    draw_focal_point  = FALSE
  )

  # C2
  log_C2 <- make_maps(
    focal_ll = mir_pts_ll, focal_m = mir_pts_m, focal_id_vec = mir_ids,
    id_name = "ID",
    out_subdir = "MIR_context10_hab10",
    hab_km = 20, context_km = 10,
    include_inset = TRUE,
    label_sfer_points = TRUE,
    label_mir_sites   = TRUE,
    mir_name_field    = "MIRSite",
    draw_focal_point  = FALSE
  )
}

# -----------------------------
# 9) Write combined log
# -----------------------------
log_all <- dplyr::bind_rows(log_A, log_B, log_C, log_C2)
readr::write_csv(log_all, file.path(run_dir, "Tables", "map_outputs_log.csv"))

message("Done. Outputs saved to: ", normalizePath(run_dir, winslash = "/", mustWork = FALSE))
```


```{r map-output-helpers, echo=FALSE, include=FALSE}
# ---- helpers ----

# Extract "ctx##km" and turn into "scale ## km"
extract_scale_from_filename <- function(fn) {
  m <- regexpr("_ctx[0-9]+km_", fn)
  if (m[1] == -1) return(NA_character_)
  hit <- regmatches(fn, m)            # like "_ctx30km_"
  km  <- gsub("[^0-9]", "", hit)
  paste0("scale ", km, " km")
}

# Clean filename -> title (SFER)
# NOTE: no "SFER" prefix here
sfer_title_fun <- function(fn) {
  site_id <- gsub(".*_ctx[0-9]+km_([^/]+)\\.png$", "\\1", fn)
  sc <- extract_scale_from_filename(fn)
  if (!is.na(sc)) paste0(site_id, " (", sc, ")") else site_id
}

# Clean filename -> title (MIR)
# NOTE: no "MIR Site:" prefix here
mir_title_fun <- function(fn) {
  mir_site <- gsub(".*_ctx[0-9]+km_([^/]+)\\.png$", "\\1", fn)
  mir_site <- gsub("_", " ", mir_site)

  # replace short codes with full names
  for (abbr in names(mir_name_lookup)) {
    mir_site <- gsub(
      paste0("\\b", abbr, "\\b"),
      mir_name_lookup[[abbr]],
      mir_site
    )
  }

  sc <- extract_scale_from_filename(fn)
  if (!is.na(sc)) paste0(mir_site, " (", sc, ")") else mir_site
}


# Render all PNGs in a folder as a list (click-to-zoom via GLightbox)
render_map_list <- function(subdir, title_fun, gallery_id = "maps", order_vec = NULL) {
  if (!dir.exists(subdir)) {
    cat(sprintf("<p><em>Folder not found:</em> %s</p>", subdir))
    return(invisible(NULL))
  }

  pngs <- list.files(subdir, pattern="\\.png$", full.names=FALSE)
  if (length(pngs) == 0) {
    cat("<p><em>No maps found.</em></p>")
    return(invisible(NULL))
  }

  # ---- ORDER CONTROL ----
  if (!is.null(order_vec)) {
    # extract ID part from filename: everything after _ctx##km_ and before .png
    ids <- gsub(".*_ctx[0-9]+km_([^/]+)\\.png$", "\\1", pngs)

    ord <- match(ids, order_vec)                # desired positions
    ord[is.na(ord)] <- 1e9 + seq_len(sum(is.na(ord)))  # shove unknowns to end, keep stable
    pngs <- pngs[order(ord, ids)]
  } else {
    pngs <- sort(pngs)
  }

  for (fn in pngs) {
    rel_fs  <- file.path(subdir, fn)   # filesystem path
    rel     <- to_web(rel_fs)          # web path for GitHub Pages

    title <- title_fun(fn)

    cat(sprintf("<h4 class='toc-ignore'>%s</h4>\n", title))
    cat(sprintf(
      "<a href='%s' class='glightbox' data-gallery='%s'>
         <img src='%s' style='width:220px; height:auto; border:1px solid #ddd; border-radius:6px; margin:6px 0;'/>
       </a>\n",
      rel, gallery_id, rel
    ))
    cat("<hr/>\n")
  }

  invisible(NULL)
}
```



## Map Outputs
<details><summary><strong>SFER — scale 30 km (`r length(list.files(file.path(run_dir,"Figures","SFER_context30_hab50"), pattern="\\.png$"))` maps)</strong></summary>
```{r, echo=FALSE, results='asis', comment=NA}
subdir <- file.path(run_dir, "Figures", "SFER_context30_hab50")
render_map_list(subdir, sfer_title_fun, gallery_id = "sfer30", order_vec = sfer_order)
```
</details>

<details><summary><strong>SFER — scale 10 km (`r length(list.files(file.path(run_dir,"Figures","SFER_context10_hab25_zoom"), pattern="\\.png$"))` maps)</strong></summary>
```{r, echo=FALSE, results='asis', comment=NA}
subdir <- file.path(run_dir, "Figures", "SFER_context10_hab25_zoom")
render_map_list(subdir, sfer_title_fun, gallery_id = "sfer10", order_vec = sfer_order)
```
</details>

<details><summary><strong>MIR — scale 10 km (`r length(list.files(file.path(run_dir,"Figures","MIR_context10_hab10"), pattern="\\.png$"))` maps)</strong></summary>
```{r, echo=FALSE, results='asis', comment=NA}
subdir <- file.path(run_dir, "Figures", "MIR_context10_hab10")
render_map_list(subdir, mir_title_fun, gallery_id = "mir10", order_vec = mir_order)
```
</details>

<details><summary><strong>MIR — scale 5 km (`r length(list.files(file.path(run_dir,"Figures","MIR_context5_hab10"), pattern="\\.png$"))` maps)</strong></summary>
```{r, echo=FALSE, results='asis', comment=NA}
subdir <- file.path(run_dir, "Figures", "MIR_context5_hab10")
render_map_list(subdir, mir_title_fun, gallery_id = "mir5", order_vec = mir_order)
```
</details>


## Proximity to SFER 
```{r mir-nearest-table-dt, echo=FALSE, results='asis'}
# nearest_tbl_path <- "/Users/heidi.k.hirsh/Desktop/MIRmodeling/tables/MIR_nearest4_SFER_20260208_081529.csv"
# nearest_tbl_path <- file.path("tables", "MIR_nearest4_SFER_20260208_081529.csv")
nearest_tbl_path <- file.path(run_dir, "Tables", "MIR_nearest4_SFER_20260208_081529.csv")


# cat("<p><b>Nearest-table file exists?</b> ", file.exists(nearest_tbl_path), "</p>")
# cat("<p><b>Working dir:</b> ", getwd(), "</p>")

cat("<h3>Closest SFER sites to each MIR site</h3>")

if (!file.exists(nearest_tbl_path)) {
  cat("<p><b>File not found:</b> ", nearest_tbl_path, "</p>", sep = "")
} else {

  nearest_tbl <- readr::read_csv(nearest_tbl_path, show_col_types = FALSE)

  # round dist_km if it exists
  if ("dist_km" %in% names(nearest_tbl)) {
    nearest_tbl <- dplyr::mutate(nearest_tbl, dist_km = round(dist_km, 2))
  }

  DT::datatable(
    nearest_tbl,
    rownames = FALSE,
    filter = "top",
    options = list(
      pageLength = 12,
      scrollX = TRUE,
      initComplete = DT::JS(
        "function(){",
        "  var api = this.api();",
        "  var col = 0; // first column = MIR site",
        "  var cell = $(api.table().header()).find('tr:eq(1) th').eq(col);",
        "  cell.empty();",
        "  var select = $('<select style=\"width: 100%\"><option value=\"\">Select MIR site…</option></select>')",
        "    .appendTo(cell)",
        "    .on('change', function(){",
        "      var v = $.fn.dataTable.util.escapeRegex($(this).val());",
        "      api.column(col).search(v ? '^'+v+'$' : '', true, false).draw();",
        "    });",
        "  api.column(col).data().unique().sort().each(function(d){",
        "    select.append('<option value=\"'+d+'\">'+d+'</option>');",
        "  });",
        "  // default: show the first MIR site only",
        "  var first = api.column(col).data()[0];",
        "  if(first !== undefined){ select.val(first).trigger('change'); }",
        "}"
      )
    ),
    caption = htmltools::tags$caption("Use the dropdown to view one MIR site at a time.")
  )
}
```

